---
title: 'Analyzing data'
author: "Statistical Programming with R"
date: ""
output:
  ioslides_presentation:
    fig_height: 5
    fig_width: 8
    logo: logo_en.png
    transition: faster
    smaller: true
  beamer_presentation: default
---

```{r, show = FALSE}
load("Contents/Material/Part C - Data manipulation/cpi_sample.RData")

```

# Packages we use

```{r}
library(dplyr)
library(magrittr)
library(tidyselect)
library(tidyverse)
library(datasets)
```


# This lecture
- Renaming data
- Selecting variables
- Aggregating data
- Joins

# Renaming data

## Simplest way
```{r}
var1 <- c(1,2,4,6,8,10)
var1
var2 <- var1
var2

```
We will return a bit to renaming later.

# Selections

## Selecting variables with [*] (Part 1)
Basic R comes with the [] operator, which can be used to select elements in vectors and lists, and columns rows from dataframes or matrices.

- For a vector `my_vect`, the command `my_vect[n]` returns the n’th element.

- For a dataframe `my_df`, the command `my_df[n,m]` returns the n’th element of the m’th column

- Writing `df[n,]` returns the n’th row

- What do you think the command `my_df[,m]` returns?

## Selecting with [*] (Part 2)
- It is possible to select by using names, too:
`my_df[c("rowname 1", "rowname 2"), c("columnname 1", "columnname 2", "columnname 3")]`

will return a subset of the dataframe `my_df` consisting of the intersection of the 2 rows named `rowname 1` and `rowname 2` and the 3 columns named `columnname 1`, `columnname 2` and `columnname 3`.

- which dimensions would that subset have?

- for lists, we use double brackets: [[]]


## More than one element
- Several elements of the same type, e.g. the integers from 1 to 5, can be concatenated into a vector
```{r}
a <- c(1, 2, 3, 4, 5)
a
```

- As is typical for `R` there a several ways to accomplish this
```{r}
b <- 1:5
b
```