---
title: "Practical C"
author: "Statistical Programming in R"
date: ""
output: 
html_document:
self_contained: false
lib_dir: libs
---


```{r echo=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  eval    = FALSE
)

```

---

#### Exercises
In this practical we are going to use sample CPI data to practice with data manipulation.


---

The following packages are required for this practical:
```{r}
library(dplyr)
library(magrittr)
```

---

1. **Open the workspace cpi_sample.RData. You need to download this file and put it in the project folder.**


---

2. **Inspect the data, how many rows and columns are there in the data? What does that mean?**

---

---

3. **Further inspect the data, what does this data set show? What can this data be used for?**

---

4. **In the `cpi_sample` data set, ..... Use a pipe to transform `hgt` in the `boys` dataset to height in meters and verify the transformation.**

<!-- Using the standard `%>%` and the `%$%` pipes: -->
```{r, echo = F}
boys %>%
  transform(hgt = hgt / 100) %$%
  mean(hgt, na.rm = TRUE)
```

---

End of Practical

---
