---
title: "Practical C"
author: "Statistical Programming in R"
date: ""
output: 
html_document:
self_contained: false
lib_dir: libs
---


```{r echo=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  eval    = FALSE
)

```

---

#### Exercises
In this practical we are going to use ITAS/ECR/Asycuda data to practice with data manipulation.


---

The following packages are required for this practical:
```{r}
library(dplyr)
library(magrittr)
library(readxl)
library(readr)
```

---

1. **Open the dataset that you've brought with you. You should first put this file in the project folder.**

```{r}
# hint -- readr or readxl may be what you need depending on the format
```
---

2. **Inspect the data: (1) how many rows and columns are there in the data? (2) What are the data types of the different variables?**

```{r}
# hint -- look at functions dim and str 
```
---

3. **Create two datasets containing rows matching certain conditions, once using subscripting and again using dplyr:filter.**

```{r}
# hint -- in the console, type: ?filter

```
---

4. **Create new datasets that contain only three columns (of your choice), once using subscripting and again using dplyr:select.** 
```{r}
# hint -- in the console, type ?select
```
---

5. **Rename each variable in your smaller datasets, once using colnames() and again using dplyr:rename.**
```{r}
#hint -- look up rename
```
---

6. **Write a single command that reads, filters, selects columns, and renames the variable using pipes.**
```{r}
# hint -- pipes are constructed using the %>% operator
```

---

End of Practical

---
