---
title: "Practical C"
author: "Statistical Programming in R"
date: ""
output: 
html_document:
self_contained: false
lib_dir: libs
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  eval    = FALSE
)

```

---

#### Exercises
In this practical we are going to use sample CPI data to practice with data manipulation.


---


1.**Open the workspace cpi_sample.RData. You need to download this file and put it in the project folder.**
```{r}
getwd()
load("cpi_sample.rdata")
summary(cpi_sample)
```
---

2. **Inspect the data, how many rows and columns are there in the data? What does that mean?**
```{r}
nrow(cpi_sample)
ncol(cpi_sample)
dim(cpi_sample)

```
---

3. **Further inspect the data, what does this data set show? What can this data be used for?**
```{r}
glimpse(cpi_sample)
str(cpi_sample)
head(cpi_sample, 10)
tail(cpi_sample, 10)
cpi_sample
View(cpi_sample)

```

---
4. **In the `cpi_sample` data set, use a pipe and the select function to get prices and quantities from January to March.**
```{r}
a <- cpi_sample%>%
  subset(select = c(PR_0121, QTY_0121, PR_0221, QTY_0221, PR_0321, QTY_0321)) %>% 
  dim()
  
b <- cpi_sample%>%
  select(PR_0121, QTY_0121, PR_0221, QTY_0221, PR_0321, QTY_0321) %>% 
  dim()
```
---
5. **Following from question 4,  Use a pipe to generate a new variable `PR_0121_new`, which is twice the old variable `PR_0121`.**

```{r}
cpi_sample%>%
  subset(select = c(PR_0121, QTY_0121, PR_0221, QTY_0221, PR_0321, QTY_0321))%>%
  mutate(PR_0121_new=PR_0121*2)%>%
  head()
```

---
6. **Following from question 4, also use the transform function to transform the column `PR_0121` into half of it.**
```{r}
cpi_sample%>%
  subset(select = c(PR_0121, QTY_0121, PR_0221, QTY_0221, PR_0321, QTY_0321))%>%
  transform(PR_0121_new = PR_0121/2) %>% 
  head()
  
```

---

7. **Filter the data for only `Greater Accra` region. Also filter out the missing responses in `outlet_type` variable**

```{r}
c <- cpi_sample%>%
  filter(!is.na(Outlet_type), 
         Region_Name=="Greater Accra") %>% 
        head()
```
---

8. **Filter the data for prices in January greater than 5**

```{r}  

cpi <- cpi_sample%>%
    filter(PR_0121>5)
``` 
--- 

9. **Rename `PR_0121` as price_jan_21**

```{r} 
library(tidyverse)

z <- cpi_sample %>%
    rename(price_jan_21=PR_0121)

```
--- 

10. **Arrange the data by PR_0121 **

```{r} 
cpi_sample %>%
  arrange(PR_0121)
```
--- 
11. **Arrange the data by PR_0121 in descending order **
---
```{r} 
cpi_sample %>%
  arrange( desc(PR_0121))

```
--- 



End of Practical

---
